---
title: "Explore Area Diff in Catchments used for LandUse Proportion Computtions"
author: "Margaux"
date: "3/31/2022"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r, message = FALSE, results='hide'}

library(dplyr)
library(tidyverse)
library(nhdplusTools)
library(mapview)
library(sf)

```

```{r,message = FALSE, results='hide'}

# Comparing PRMS NLCD proportions dataset wih FORESCE
targets::tar_make(p2_PRMS_NLCD_lc_proportions_reclass_cat)
targets::tar_make(p2_FORESCE_LC_per_catchment_reclass_cat)
targets::tar_load(p2_PRMS_NLCD_lc_proportions_reclass_cat)
targets::tar_load(p2_FORESCE_LC_per_catchment_reclass_cat)

targets::tar_make(p1_catchments_edited_sf)
targets::tar_load(p1_catchments_edited_sf)
targets::tar_make(p1_reaches_sf)
targets::tar_load(p1_reaches_sf)

```

### Compare NLCD and FORESCE area in lu proportion table 

```{r}

# subset/filter and mutate. We choose year 2000 for FORESCE and 2001 for NLCD

##2000
l1 <- p2_FORESCE_LC_per_catchment_reclass_cat[[7]] %>%
  #  select(PRMS_segid, Year, everything(), -hru_segment) %>% 
  mutate(total_PRMS_area_km2 = total_PRMS_area/10^6) %>%
  select(PRMS_segid, hru_segment, total_PRMS_area_km2)
## 2001
l2 <- p2_PRMS_NLCD_lc_proportions_reclass_cat[[1]] %>%
  # select(PRMS_segid, Year, everything(), -AREASQKM_PRMS) %>% 
  select(PRMS_segid, AREASQKM_PRMS)

head(l1)
head(l2)

```

```{r}
# Merge l1 and l2 and calculate different for each PRMS SEGMENT
df <- l1 %>% left_join(l2, by = 'PRMS_segid') %>% 
  rename(PRMS_area_km2_FOR = total_PRMS_area_km2, PRMS_area_km2_NLCD = AREASQKM_PRMS) %>% 
  mutate(ab_diff = abs(PRMS_area_km2_FOR - PRMS_area_km2_NLCD)) %>%
  mutate(ab_diff_categorized = factor(case_when(ab_diff > 50 ~ '>50 km2',
                              ab_diff > 10 ~ '>10 km2',
                              ab_diff > 5 ~ '>5 km2',
                              ab_diff >1 ~ '>1 km2',
                              TRUE ~ '< 1 km2'),
                              levels = c('< 1 km2', '>1 km2', '>5 km2', '>10 km2', '>50 km2'))) 


dim(df)

```

```{r}

## Plotting
scatterplot1 <- ggplot(df, aes(y = PRMS_area_km2_FOR, x = PRMS_area_km2_NLCD))+
  geom_point(aes(color = ab_diff_categorized))+theme_bw()+
  ylab('PRMS Cat Area - geopackage')+
  xlab('PRMS Cat Area - nhd-comid catchment aggregation')+
  labs(caption = 'Scatter plot of PRMS area in the NLCD dataset versus FORESCE.\n Point colored by their difference categories')+
  theme(plot.caption = element_text(hjust = 0.5))
scatterplot1  

```

```{r}
bar_plot_diff_bins <- ggplot(df, aes(x = ab_diff_categorized))+
  geom_bar(fill = 'lightblue4')+
  scale_y_continuous(n.breaks = 15)+
  theme_bw()+
  labs(caption ='\n Ab difference between area (km2) of FORESCE land use prop dataset\n (2000) and NLCD land use dataset (2001) grouped into 4 categories')+
  theme(plot.caption = element_text(hjust = 0.5))
bar_plot_diff_bins

```

### Mapping errorenous PRMS segments
```{r}
## Extract rows that are problematic
over_50_km2 <- df %>% filter(ab_diff_categorized == '>50 km2') %>% select(PRMS_segid)
over_10_km2 <- df %>% filter(ab_diff_categorized == '>10 km2') %>% select(PRMS_segid)
over_5_km2 <- df %>% filter(ab_diff_categorized == '>5 km2') %>% select(PRMS_segid)
over_1_km2 <- df %>% filter(ab_diff_categorized == '>1 km2') %>% select(PRMS_segid)
under_1_km2 <- df %>% filter(ab_diff_categorized == '< 1 km2') %>% select(PRMS_segid)

## Pulling target that maps all comids within each PRMS
targets::tar_make(p2_drb_comids_all_tribs)
targets::tar_load(p2_drb_comids_all_tribs)

## subsetting to the comids that have differing areas
comids_over_50_km2 <- p2_drb_comids_all_tribs %>%
  filter(PRMS_segid %in% over_50_km2$PRMS_segid)

comids_over_10_km2 <- p2_drb_comids_all_tribs %>%
  filter(PRMS_segid %in% over_10_km2$PRMS_segid)

## subsetting to the comids that have differing areas
comids_over_5_km2 <- p2_drb_comids_all_tribs %>%
  filter(PRMS_segid %in% over_5_km2$PRMS_segid)

comids_over_1_km2 <- p2_drb_comids_all_tribs %>%
  filter(PRMS_segid %in% over_1_km2$PRMS_segid)

comids_under_1_km2 <- p2_drb_comids_all_tribs %>%
  filter(PRMS_segid %in% under_1_km2$PRMS_segid)


head(comids_over_50_km2)

```


```{r, message = FALSE, results='hide'}
# Check whether incorrect comids originally had comids of area 0
targets::tar_make(p2_NLCD_LC_w_catchment_area)
targets::tar_load(p2_NLCD_LC_w_catchment_area)

```

```{r}

## Subset the source dataset for NLCD w/ area 
PRMS_areasqkm_0 <- p2_NLCD_LC_w_catchment_area %>%
  filter(AREASQKM == 0) %>% select(PRMS_segid, comid, AREASQKM)
PRMS_areasqkm_0

# Print list of PRMS that originally had areasqkm of 0 and that got an total_area val of lengthkm^2 
PRMS_areasqkm_0 %>% 
       mutate(incorrect = case_when(PRMS_segid %in%
                                      comids_over_50_km2$PRMS_segid ~ "TRUE",
                                    TRUE ~ 'FALSE')) %>% 
  filter(incorrect == TRUE)
     
```

```{r}

map_difference<- function(df){
  
  catchment_gpkg <- p1_catchments_edited_sf %>% filter(PRMS_segid %in% unique(df$PRMS_segid))
  reach <- p1_reaches_sf %>% filter(subsegid %in% unique(df$PRMS_segid))
  
  mapview(df, col.regions = 'red', alpha.regions = 0.7)+
    mapview(catchment_gpkg, alpha.regions = 0.5)+
    mapview(reach, color = 'black')

}

```

## Catchments with over 50km2 diff
```{r}

## Grab comids with over 50 km2 difference using nhdplustools
all_comids <- get_nhdplus(comid = comids_over_50_km2$comid, realization = 'catchment') %>% 
  mutate(featureid = as.character(featureid))

## join with our comids_over_50km2 to have prms col
comids_over_50_km2 <- comids_over_50_km2 %>%
  left_join(all_comids, by= c('comid' = 'featureid'),
            keep = TRUE) %>% 
  ## make sf object
  sf::st_as_sf()

# sample_n(comids_over_50_km2,5)

map_difference(comids_over_50_km2)

```

## Catchments with over 10km2 diff
```{r}

## Grab comids with over 10 km2 difference using nhdplustools
all_comids <- get_nhdplus(comid = comids_over_10_km2$comid, realization = 'catchment') %>% 
  mutate(featureid = as.character(featureid))

## join with our comids_over_50km2 to have prms col
comids_over_10_km2 <- comids_over_10_km2 %>%
  left_join(all_comids, by= c('comid' = 'featureid'),
            keep = TRUE) %>% st_as_sf()

sample_n(comids_over_10_km2, 10)

map_difference(comids_over_10_km2)

```
### >5 km2 - 10 km2

```{r}

## Grab comids with over 5 km2 difference using nhdplustools
all_comids <- get_nhdplus(comid = comids_over_5_km2$comid, realization = 'catchment') %>%
  mutate(featureid = as.character(featureid))

## join with our comids_over_50km2 to have prms col
comids_over_5_km2 <- comids_over_5_km2 %>%
  left_join(all_comids, by= c('comid' = 'featureid'),
            keep = TRUE) 

sample_n(comids_over_5_km2,10)

map_difference(comids_over_5_km2)

```

### > 1km2 - <5 km2
```{r}

## Grab comids with over 1 km2 difference using nhdplustools
all_comids <- get_nhdplus(comid = comids_over_1_km2$comid, realization = 'catchment') %>%
  mutate(featureid = as.character(featureid))

## join with our comids_over_1km2 to have prms col
comids_over_1_km2 <- comids_over_1_km2 %>%
  left_join(all_comids, by= c('comid' = 'featureid'),
            keep = TRUE) %>% st_as_sf()

sample_n(comids_over_1_km2,10)

map_difference(comids_over_1_km2)


```

### <1 km2 diff
```{r}
# 
# ## Grab comids with over 1 km2 and less than km 2 - difference using nhdplustools
# all_comids <- get_nhdplus(comid = comids_under_1_km2$comid[1:10], realization = 'catchment') %>% 
#   mutate(featureid = as.character(featureid))
# 
# ## join with our comids_over_1km2 to have prms col
# comids_under_1_km2 <- comids_under_1_km2 %>%
#   left_join(all_comids, by= c('comid' = 'featureid'),
#             keep = TRUE) %>% sf::st_as_sf()
# 
# 
# map_difference(comids_under_1_km2)
# 

```


### Examining specific catchments area with diff > 50 kms
```{r}

## Assess specific PRMS_segid - 12_1
prms_seg <- '12_1'
print(df[,1:3][df$PRMS_segid == prms_seg,])

subsetted_comids_over_50km2 <- comids_over_50_km2 %>% filter(PRMS_segid == prms_seg)
catchment <- p1_catchments_edited_sf %>% filter(PRMS_segid == unique(subsetted_comids_over_50km2$PRMS_segid))
reach <- p1_reaches_sf %>% filter(subsegid == unique(subsetted_comids_over_50km2$PRMS_segid))
                                  
mapview(subsetted_comids_over_50km2,
        col.regions = 'red', alpha.regions = 0.5) +
  mapview(catchment)+ mapview(reach, color = 'black')

```

```{r}
## 157_1
prms_seg <- '157_1'
print(df[,1:3][df$PRMS_segid == prms_seg,])

subsetted_comids_over_50km2 <- comids_over_50_km2 %>% filter(PRMS_segid == prms_seg)
catchment <- p1_catchments_edited_sf %>% filter(PRMS_segid == unique(subsetted_comids_over_50km2$PRMS_segid))
reach <- p1_reaches_sf %>% filter(subsegid == unique(subsetted_comids_over_50km2$PRMS_segid))

mapview(subsetted_comids_over_50km2,
        col.regions = 'red', alpha.regions = 0.5) +
  mapview(catchment)+ mapview(reach, color = 'black')

comids_over_50_km2

```

```{r}
## 216_1
prms_seg <- '216_1'
print(df[,1:3][df$PRMS_segid == prms_seg,])

subsetted_comids_over_50km2 <- comids_over_50_km2 %>% filter(PRMS_segid == prms_seg)
catchment <- p1_catchments_edited_sf %>% filter(PRMS_segid == unique(subsetted_comids_over_50km2$PRMS_segid))
reach <- p1_reaches_sf %>% filter(subsegid == unique(subsetted_comids_over_50km2$PRMS_segid))

mapview(subsetted_comids_over_50km2,
        col.regions = 'red', alpha.regions = 0.5) +
  mapview(catchment)+ mapview(reach, color = 'black')

```

```{r}
## 878_1
prms_seg <- '878_1'
print(df[,1:3][df$PRMS_segid == prms_seg,])

subsetted_comids_over_50km2 <- comids_over_50_km2 %>% filter(PRMS_segid == prms_seg)
catchment <- p1_catchments_edited_sf %>% filter(PRMS_segid == unique(subsetted_comids_over_50km2$PRMS_segid))
reach <- p1_reaches_sf %>% filter(subsegid == unique(subsetted_comids_over_50km2$PRMS_segid))

mapview(subsetted_comids_over_50km2,
        col.regions = 'red', alpha.regions = 0.5) +
  mapview(catchment)+ mapview(reach, color = 'black')
```

```{r}

## 386_1
prms_seg <- '386_1'
print(df[,1:3][df$PRMS_segid == prms_seg,])

subsetted_comids_over_50km2 <- comids_over_50_km2 %>% filter(PRMS_segid == prms_seg)
catchment <- p1_catchments_edited_sf %>% filter(PRMS_segid == unique(subsetted_comids_over_50km2$PRMS_segid))
reach <- p1_reaches_sf %>% filter(subsegid == unique(subsetted_comids_over_50km2$PRMS_segid))

mapview(subsetted_comids_over_50km2,
        col.regions = 'red', alpha.regions = 0.5) +
  mapview(catchment)+ mapview(reach, color = 'black')

```

```{r}

comids_386_1 <- p2_drb_comids_all_tribs %>% filter(PRMS_segid =='386_1')
comids_387_1 <- p2_drb_comids_all_tribs %>% filter(PRMS_segid =='387_1')

## Polygon is the same 
PRMS_catchments_386_387 <- p1_catchments_edited_sf[p1_catchments_edited_sf$PRMS_segid %in% c('386_1','387_1'),]

mapview(get_nhdplus(comid = comids_386_1$comid, realization = 'catchment'))+
  mapview(get_nhdplus(comid = comids_387_1$comid, realization = 'catchment'), col.regions = 'red')+
  mapview(PRMS_catchments_386_387, col.regions = 'green')
```


### Explore the geopackage

```{r}

## catchment for prms >50 km2 diff overlay on all catchments 
mapview(p1_catchments_edited_sf %>% filter(PRMS_segid %in% over_50_km2$PRMS_segid), color = 'blue', col.regions = 'transparent', lwd = 2)+
  mapview(p1_catchments_edited_sf, col.regions = 'red', alpha.regions = 0.3)

```

### Checking area diff between the upstream targets - p1_catchments_edited_sf vs p2_NLCD_LC_w_catchment_area

```{r}

# Check areas of upstream targets used to generate the land use targets above. 

targets::tar_load(p1_catchments_edited_sf)
targets::tar_load(p2_NLCD_LC_w_catchment_area)

t1 <- p2_NLCD_LC_w_catchment_area %>% group_by(PRMS_segid) %>% 
  summarise(PRMS_area_km2 = sum(AREASQKM))

t2 <- p1_catchments_edited_sf %>% sf::st_drop_geometry() %>% 
  group_by(PRMS_segid) %>% summarise(PRMS_area_km2 = sum(hru_area_m2)/10^6)


df_2 <- t1 %>% left_join(t2, by = 'PRMS_segid') %>% 
  mutate(ab_diff = abs(PRMS_area_km2.x - PRMS_area_km2.y)) %>%
  mutate(ab_diff_categorized = factor(case_when(ab_diff > 50 ~ '>50 km2',
                                                ab_diff > 10 ~ '>10 km2',
                                                ab_diff > 5 ~ '>5 km2',
                                                ab_diff >1 ~ '>1 km2',
                                                TRUE ~ '< 1 km2'),
                                      levels = c('< 1 km2', '>1 km2', '>5 km2', '>10 km2', '>50 km2'))) 

ggplot(df_2, aes(x = ab_diff_categorized))+
  geom_bar()+
  scale_y_continuous(n.breaks = 15)+theme_bw()+
  labs(caption ='\n Ab difference between area (km2) of catchment polygons in p1_catchments_edited_sf and xwalk area from land use prop dataset\n (2000) and NLCD land use dataset (2001) grouped into 4 categories')

```


the same 

--TBD--

### Looking at old shp file to see what PRMS segs previously had no catchments
```{r}

# Download PRMS catchments for region 02
  ## Downloaded from ScienceBase: https://www.sciencebase.gov/catalog/item/5362b683e4b0c409c6289bf6
source("1_fetch/src/get_gf.R")
# Define dataset of interest for the national geospatial fabric (used to fetch PRMS catchment polygons):
gf_data_select <- 'GeospatialFabricFeatures_02.zip'
p1_catchments_shp <- get_gf(out_dir = "1_fetch/out/", sb_id = '5362b683e4b0c409c6289bf6', sb_name = gf_data_select)
  
# Read PRMS catchment shapefile into sf object and filter to DRB
p1_catchments_sf<- sf::st_read(dsn = p1_catchments_shp,layer="nhru", quiet=TRUE) %>% 
  filter(hru_segment %in% p1_reaches_sf$subsegseg) %>%
  suppressWarnings()
  
## Fix issue geometries in p1_catchments_sf by defining a 0 buffer around polylines
p1_catchments_sf_valid <- sf::st_buffer(p1_catchments_sf,0)
  

```
